---
title: "HW10"
author: "Tracie-Lynn Lamoreux"
date: "2023-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(shiny)
library(maps)
acc <- read.csv("https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/accident.csv", stringsAsFactors = FALSE)
```

```{r}
states <- map_data("state")
head(states)
acc
names(acc)[names(acc) == "LONGITUD"] <- "long"
names(acc)[names(acc) == "LATITUDE"] <- "lat"
us_acc <- acc[acc$lat >= 24.5 & acc$lat <= 49.5 & 
                acc$long >= -124.8 & acc$long <= -66.9, ]
map <- states %>% ggplot(aes(x = long, y = lat)) + geom_path(aes(group = group)) + geom_point(data = us_acc, aes(x=long, y = lat))
```


```{r}
day_freq <- table(acc$DAY_WEEK)
day_freq
acc2 <- acc %>%  filter(acc$HOUR < 30) 
ggplot(acc2, aes(x = HOUR)) +
  geom_histogram(binwidth = 1, fill = "blue") +
  ggtitle("Accidents by Hour of the Day") +
  xlab("Hour of the Day") +
  ylab("Number of Accidents")
```

Accidents are more common on days 1, 6, and 7 (presumably the three weekend days). Day 7 has almost 1000 more accidents than day 3. Similar, accidents are more likely to occur during the evening, peaking at 18:00-19:00. The time with the least accidents is early morning, from 4:00-5:00. There is a brief peak at 6:00-7:00 as people go to work, it drops in the early morning, then climbs to the evening.

```{r}
acc$DRUNK_DR<- as.numeric(acc$DRUNK_DR)
atleast1 <- acc %>%
  filter(DRUNK_DR >= 1)
nrow(atleast1)
```

There are 8720 accidents with at least one drunk driver.

```{r}
person <- read.csv("https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/person.csv", stringsAsFactors = FALSE)
drivers <- person %>%
  filter(PER_TYP == 1)
accident_with_drivers <- acc %>%
  inner_join(drivers, by = "ST_CASE")
```


```{r}
accidents_by_day_hour_gender <- accident_with_drivers %>%
  group_by(DAY_WEEK, HOUR.x, SEX) %>%
  summarize(n = n())

accidents_by_day_hour_gender <- accidents_by_day_hour_gender %>% filter(HOUR.x <30)

```

```{r}
#making SEX into a factor so it separates out each line by gender--the graph was weird, nearly unreadable without doing this
accidents_by_day_hour_gender <- accidents_by_day_hour_gender %>% mutate(SEX = as.factor(SEX))
```


```{r}

ggplot(accidents_by_day_hour_gender, aes(x = HOUR.x, y = n, color = SEX)) +
  geom_line(size = 1) +
  facet_wrap(~DAY_WEEK, ncol = 3) +
  labs(x = "Hour of the day", y = "Number of accidents", color = "Gender")
```